---
title: "Business Intelligence Project"
author: "korn"
date: "23/10/23"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

# Student Details

+-------------------------------------+--------------------------------+
| **Student ID Numbers and Names of   | 1.  134644 - C - Sebastian     |
| Group Members**                     |     Muramara                   |
|                                     |                                |
|                                     | 2.  136675 - C - Bernard       |
|                                     |     Otieno                     |
|                                     |                                |
|                                     | 3.  131589 - C - Agnes Anyango |
|                                     |                                |
|                                     | 4.  131582 - C - Njeri Njuguna |
|                                     |                                |
|                                     | 5.  136009 - C- Sera Ndabari   |
+-------------------------------------+--------------------------------+
| **GitHub Classroom Group Name**     | Korn                           |
+-------------------------------------+--------------------------------+
| **Course Code**                     | BBT4206                        |
+-------------------------------------+--------------------------------+
| **Course Name**                     | Business Intelligence II       |
+-------------------------------------+--------------------------------+
| **Program**                         | Bachelor of Business           |
|                                     | Information Technology         |
+-------------------------------------+--------------------------------+
| **Semester Duration**               | 21^st^ August 2023 to 28^th^   |
|                                     | November 2023                  |
+-------------------------------------+--------------------------------+

# Setup Chunk

**Note:** the following KnitR options have been set as the global
defaults: <BR>
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here
<https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and
here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)

```

# LAB 7

## CLASSIFICATION

### Load required Packages

```{r Packages}

if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## stats ----
if (require("stats")) {
  require("stats")
} else {
  install.packages("stats", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## MASS ----
if (require("MASS")) {
  require("MASS")
} else {
  install.packages("MASS", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## glmnet ----
if (require("glmnet")) {
  require("glmnet")
} else {
  install.packages("glmnet", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071 ----
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## kernlab ----
if (require("kernlab")) {
  require("kernlab")
} else {
  install.packages("kernlab", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## rpart ----
if (require("rpart")) {
  require("rpart")
} else {
  install.packages("rpart", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

```

#Dataset used for classification

## Dataset Description

A data frame with 214 observation containing examples of the chemical
analysis of 7 different types of glass. The problem is to forecast the
type of class on basis of the chemical analysis. The study of
classification of types of glass was motivated by criminological
investigation. At the scene of the crime, the glass left can be used as
evidence (if it is correctlyidentified!).

#Usage

data(Glass)

#Format

A data frame with 214 observations on 10 variables:

[,1] RI refractive index

[,2] Na Sodium

[,3] Mg Magnesium

[,4] Al Aluminum

[,5] Si Silicon

[,6] K Potassium

[,7] Ca Calcium

[,8] Ba Barium

[,9] Fe Iron

[,10] Type Type of glass (class attribute)

#Source

• Creator: B. German, Central Research Establishment, Home Office
Forensic Science Service, Aldermaston, Reading, Berkshire RG7 4PN

• Donor: Vina Spiehler, Ph.D., DABFT, Diagnostic Products Corporation
These data have been taken from the UCI Repository Of Machine Learning
Databases at 10 HouseVotes84

• <ftp://ftp.ics.uci.edu/pub/machine-learning-databases>

• <http://www.ics.uci.edu/~mlearn/MLRepository.html> and were converted
to R format by Friedrich Leisch.

#References

Newman, D.J. & Hettich, S. & Blake, C.L. & Merz, C.J. (1998). UCI
Repository of machine learning databases
[<http://www.ics.uci.edu/~mlearn/MLRepository.html>]. Irvine, CA:
University of California, Department of Information and Computer
Science.

#Examples

data(Glass) summary(Glass)

## 1.a.Decision tree for a classification problem without caret

```{r CART}
#### Load and split the dataset ----
data(Glass)

# Define a 80:20 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.8,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
Type_model_rpart <- rpart(Type ~ ., data = Glass_train)

#### Display the model's details ----
print(Type_model_rpart)

#### Make predictions ----
predictions <- predict(Type_model_rpart,
                       Glass_test[, 1:9],
                       type = "class")

#### Display the model's evaluation metrics ----
table(predictions,Glass_test$Type)

confusion_matrix <-
  caret::confusionMatrix(predictions,
                        Glass_test[, 1:10]$Type)
print(confusion_matrix)
```

## 1.b. Decision tree for a classification problem with caret

```{r CART with caret}
#### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.7,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
set.seed(121)
# We apply the 10-fold cross validation resampling method
train_control <- trainControl(method = "cv", number = 10)
Glass_caret_model_rpart <- train(Type ~ ., data = Glass,
                                 method = "rpart", metric = "Accuracy",
                                 trControl = train_control)

#### Display the model's details ----
print(Glass_caret_model_rpart)

#### Make predictions ----
predictions <- predict(Glass_caret_model_rpart, 
                       Glass_test[, 1:9],
                       type = "raw") 

#### Display the model's evaluation metrics ----
table(predictions, Glass_test$Type)

confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_test[, 1:10]$Type)
print(confusion_matrix)

```

## 2.a. Naïve Bayes Classifier for a Classification Problem without CARET

```{r nb}
# We use the naiveBayes function inside the e1071 package
#### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.8,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
Glass_model_nb <- naiveBayes(Type ~ .,
                                data = Glass_train)

#### Display the model's details ----
print(Glass_model_nb)

#### Make predictions ----
predictions <- predict(Glass_model_nb,
                       Glass_test[, -10])

#### Display the model's evaluation metrics ----
confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_test[, 1:10]$Type)
print(confusion_matrix)
```

## 2.b. Naïve Bayes Classifier for a Classification Problem with CARET

```{r nb with caret}
#### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.7,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
# We apply the 5-fold cross validation resampling method
set.seed(7)
train_control <- trainControl(method = "cv", number = 10)
preProcessObj <- preProcess(Glass_train, method = "medianImpute")
Glass_train_imputed <- predict(preProcessObj, Glass_train)
Glass_caret_model_nb <- train(Type ~ .,
                                 data = Glass_train_imputed,
                                 method = "nb", metric = "Accuracy",
                                 trControl = train_control)

#### Display the model's details ----
print(Glass_caret_model_nb)

#### Make predictions ----
predictions <- predict(Glass_caret_model_nb,
                       Glass_train_imputed[, 1:9])

#### Display the model's evaluation metrics ----
confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_train_imputed[, 1:10]$Type)
print(confusion_matrix)


```

## 3.a. kNN for a classification problem without CARET's train function

```{r kNN}
#### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.8,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
Glass_model_knn <- knn3(Type ~ ., data = Glass_train, k=2)

#### Display the model's details ----
print(Glass_model_knn)

#### Make predictions ----
predictions <- predict(Glass_model_knn,
                       Glass_test[, 1:9],
                       type = "class")

#### Display the model's evaluation metrics ----
table(predictions, Glass_test$Type)

# Or alternatively:
confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_test$Type)
print(confusion_matrix)


```

## 3.b. kNN for a classification problem with CARET's train function

```{r kNN with caret}
#### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.7,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----

set.seed(121)
train_control <- trainControl(method = "cv", number = 10)
Glass_caret_model_knn <- train(Type ~ ., data = Glass,
                                  method = "knn", metric = "Accuracy",
                                  preProcess = c("center", "scale"),
                                  trControl = train_control)

#### Display the model's details ----
print(Glass_caret_model_knn)

#### Make predictions ----
predictions <- predict(Glass_caret_model_knn,
                       Glass_test[, 1:9])

#### Display the model's evaluation metrics ----
confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_test[, 1:10]$Type)
print(confusion_matrix)

```

## 4.a. SVM Classifier for a classification problem with CARET

```{r SVM with caret}
 #### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.7,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
set.seed(121)
train_control <- trainControl(method = "cv", number = 11)
Glass_caret_model_svm_radial <- # nolint: object_length_linter.
  train(Type ~ ., data = Glass_train, method = "svmRadial",
        metric = "Accuracy", trControl = train_control)

#### Display the model's details ----
print(Glass_caret_model_svm_radial)

#### Make predictions ----
predictions <- predict(Glass_caret_model_svm_radial,
                       Glass_test[, 1:9])

#### Display the model's evaluation metrics ----
table(predictions, Glass_test$Type)
confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_test[, 1:10]$Type)
print(confusion_matrix)

```

## 4.b. SVM Classifier for a classification problem without CARET

```{r SVM}
#### Load and split the dataset ----
data(Glass)

# Define a 70:30 train:test data split of the dataset.
train_index <- createDataPartition(Glass$Type,
                                   p = 0.7,
                                   list = FALSE)
Glass_train <- Glass[train_index, ]
Glass_test <- Glass[-train_index, ]

#### Train the model ----
Glass_model_svm <- ksvm(Type ~ ., data = Glass_train,
                           kernel = "rbfdot")

#### Display the model's details ----
print(Glass_model_svm)

#### Make predictions ----
predictions <- predict(Glass_model_svm, Glass_test[, 1:9],
                       type = "response")

#### Display the model's evaluation metrics ----
table(predictions, Glass_test$Type)

confusion_matrix <-
  caret::confusionMatrix(predictions,
                         Glass_test[, 1:10]$Type)
print(confusion_matrix)
```
